---
title: "Interactive Graphs"
author: "Anastasiia Sviridova, Shokoufeh Naseri"
date: "2025-01-24"
output: html_document
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
Sys.setenv(LANG = "en")
options(scipen = 5)
setwd("C:/Users/User/Documents/3 semester/ad vis in r/project/R_Project")
getwd()

# Install ggpubr if not already installed
if (!requireNamespace("ggpubr", quietly = TRUE)) {
  install.packages("ggpubr")
}

# Install countrycode if not already installed
if (!requireNamespace("countrycode", quietly = TRUE)) {
  install.packages("countrycode")
}
# Install and load plotly for interactive visualizations
if (!requireNamespace("plotly", quietly = TRUE)) {
  install.packages("plotly")
}

# Load the package
library(plotly)
library(ggpubr)
library(countrycode)
library(ggplot2)
library(dplyr)
library(readr)
library(corrplot)
library(maps)
```
```{r include=FALSE}
# Load the new dataset
new_data <- read.csv("cleaned_dataset.csv")

# Subset for economic prosperity and health outcomes
data <- new_data %>%
  select(Country, GDP, MinimumWage, LifeExpectancy, InfantMortality, Latitude, Longitude)

# Subset for CO2 emissions and public health outcomes
co2_health_data <- new_data %>%
  select(Country, CO2Emissions, LifeExpectancy, InfantMortality)

# Subset for healthcare metrics and life expectancy
healthcare_data <- new_data %>%
  select(Country, PhysiciansPerThousand, HealthExpenditure, LifeExpectancy)
```

```{r include=FALSE}
# Map countries to continents using the countrycode package
data$Continent <- countrycode(data$Country, origin = "country.name", destination = "continent")


# Custom breaks for GDP axis
custom_gdp_breaks <- c(1e9, 1e10, 1e11, 1e12, 1e13)
custom_gdp_labels <- c("1B", "10B", "100B", "1T", "10T")
```


### Interactive Visualizations

These graphs allow us to explore the relationships between GDP, life expectancy, and infant mortality by hovering over data points for country-specific details.

```{r echo=FALSE}
# Create interactive visualizations
create_interactive_visualizations <- function(data) {
  # Interactive visualization for GDP and Life Expectancy
  plot1 <- plot_ly(
    data,
    x = ~GDP,
    y = ~LifeExpectancy,
    type = 'scatter',
    mode = 'markers',
    text = ~paste("Country:", Country, "<br>GDP:", GDP, "<br>Life Expectancy:", LifeExpectancy),
    marker = list(color = ~LifeExpectancy, colorscale = 'Viridis', size = ~log(GDP + 1))
  ) %>%
    layout(
      title = "Interactive: GDP vs Life Expectancy",
      xaxis = list(title = "GDP", type = "log"),
      yaxis = list(title = "Life Expectancy")
    )

  # Interactive visualization for GDP and Infant Mortality
  plot2 <- plot_ly(
    data,
    x = ~GDP,
    y = ~InfantMortality,
    type = 'scatter',
    mode = 'markers',
    text = ~paste("Country:", Country, "<br>GDP:", GDP, "<br>Infant Mortality:", InfantMortality),
    marker = list(color = ~InfantMortality, colorscale = 'Reds', size = ~log(GDP + 1))
  ) %>%
    layout(
      title = "Interactive: GDP vs Infant Mortality",
      xaxis = list(title = "GDP", type = "log"),
      yaxis = list(title = "Infant Mortality")
    )

  # Interactive visualization for Life Expectancy and Infant Mortality
  plot3 <- plot_ly(
    data,
    x = ~LifeExpectancy,
    y = ~InfantMortality,
    type = 'scatter',
    mode = 'markers',
    text = ~paste("Country:", Country, "<br>Life Expectancy:", LifeExpectancy, "<br>Infant Mortality:", InfantMortality),
    marker = list(color = ~LifeExpectancy, colorscale = 'Blues', size = ~log(GDP + 1))
  ) %>%
    layout(
      title = "Interactive: Life Expectancy vs Infant Mortality",
      xaxis = list(title = "Life Expectancy"),
      yaxis = list(title = "Infant Mortality")
    )

  # Save interactive plots as HTML
  htmlwidgets::saveWidget(plot1, "interactive_gdp_life_expectancy.html")
  htmlwidgets::saveWidget(plot2, "interactive_gdp_infant_mortality.html")
  htmlwidgets::saveWidget(plot3, "interactive_life_expectancy_infant_mortality.html")

  return(list(plot1, plot2, plot3))
}

# Generate interactive visualizations
interactive_plots <- create_interactive_visualizations(data)

# Optionally print interactive plots in RStudio Viewer
interactive_plots[[1]]
interactive_plots[[2]]
interactive_plots[[3]]


```

### Geospatial visualization for Life Expectancy

This geospatial visualization allows us to easily identify regional disparities and patterns, with richer countries generally showing higher life expectancy.


```{r echo=FALSE}
# Interactive Global Life Expectancy Map
world_map <- map_data("world") # Load world map data

interactive_map <- plot_ly(width = 1200, height = 700) %>%
  add_polygons(
    data = world_map,
    x = ~long,
    y = ~lat,
    split = ~group,
    line = list(width = 0.5),
    fillcolor = 'lightgray',
    hoverinfo = 'none'
  ) %>%
  add_markers(
    data = data,
    x = ~Longitude,
    y = ~Latitude,
    text = ~paste("Country:", Country, "<br>Life Expectancy:", LifeExpectancy),
    color = ~LifeExpectancy,
    colors = c("red", "green"),
    marker = list(size = 8, opacity = 0.7)
  ) %>%
  layout(
    title = "Interactive Global Life Expectancy",
    xaxis = list(title = "Longitude", range = c(-180, 180)),
    yaxis = list(title = "Latitude", range = c(-60, 90)),
    showlegend = FALSE
  )

# Display the interactive map
interactive_map

```


